---
title: "Empirical Analysis of AMZN Microstructure Data"
author: "Sid Bhatia"
date: "12/3/2023"
output: pdf_document
---

## Importation

Import the required libraries.

```{r}
library(highfrequency)
library(tidyverse)
library(xts)
library(lubridate)
```

## Data Pre-Processing

Load and pre-process the TAQ data.

```{r}
# Write file path for data.
file_path <- 'C://Users//sbhatia2//OneDrive - stevens.edu//Desktop//amzn_tick_history_01_04_2023.csv'

# Read in TAQ data.
data <- read.csv(file_path)

head(data)
tail(data)

# Display columns of dataset.
colnames(data)

head(data$Bid.Price)
tail(data$Bid.Price)

head(data$Ask.Price)
tail(data$Ask.Price)

# Remove rows with NA Bid/Ask/Trade prices.
cleaned_data <- data[!is.na(data$Bid.Price) & !is.na(data$Ask.Price) & !is.na(data$Price), ]

head(cleaned_data)
tail(cleaned_data)
```

## Liquidity Analysis

Perform core liquidity analysis for the project.

#### Price Computation

Retrieve and compute the bid, ask, trade, and mid prices.

```{r}
# Determine number of trades in dataset.
num_of_trades <- nrow(cleaned_data)

paste("Total number of trades for AMZN on 01/04/2023:", num_of_trades)

# Retrieve trade, bid, and ask prices.
prices <- as.numeric(cleaned_data$Price)
bids <- as.numeric(cleaned_data$Bid.Price)
asks <- as.numeric(cleaned_data$Ask.Price)

# Compute the mid price as the average of the bid and ask prices.
mids <- (bids + asks) * 0.5
```

#### Trade vs. Mid Price Plotting

Plot the trade vs. mid prices.

```{r}
plot(prices, type="l", col="red", main="Trade vs. Mid Prices", 
     xlab="Trade Count",
     ylab="Trade Price")
lines(mids, type="l", col="blue")
```

#### Size Computation

Retrieve and compute the bid and ask sizes as well as the LOB imbalances.

```{r}
# Retrieve bid and ask sizes.
bid_size <- as.numeric(cleaned_data$Bid.Size)
ask_size <- as.numeric(cleaned_data$Ask.Size)

# Compute the LOB imbalance at each time step as the sum of the bid and ask sizes.
LOB_imbalance <- bid_size + ask_size
```

#### Trade Size Plotting

Plot the trade sizes.

```{r}
plot(ask_size, col="orange", type="h", 
     ylab="Trade Size", 
     xlab="Trade Count", main="Trade Size at Best Quote")
lines(bid_size, col="purple", type="h")
```

#### Liquidity Measures

Compute the relevant liquidity measures.

```{r}
# Convert Exchange Time column to be compatible with xts object.
cleaned_data$Exch.Time <- as.POSIXct(paste("2000-01-01", cleaned_data$Exch.Time))

# Convert dataframe into xts object.
amzn_xts <- xts(cleaned_data[,-which(names(cleaned_data) == "Exch.Time")], order.by = cleaned_data$Exch.Time)

head(amzn_xts)
tail(amzn_xts)

head(amzn_xts$Bid.Price)

# Rename Bid, Ask, and Symbol columns to match correct format.
names(amzn_xts)[names(amzn_xts) == 'Bid.Price'] <- 'BID'
names(amzn_xts)[names(amzn_xts) == 'ASK'] <- 'OFR'
names(amzn_xts)[names(amzn_xts) == 'X.RIC'] <- 'SYMBOL'

# Retrieve liquidity measures.
liquidity_measures <- getLiquidityMeasures(amzn_xts)

# Compute the Quoted Spread.
quoted_spread <- mean(as.numeric(liquidity_measures$quotedSpread))

quoted_spread

# Compute the Effective Spread.
effective_spread <- mean(as.numeric(liquidity_measures$effectiveSpread))

effective_spread

# Compute the Realized Spread.
realized_spread <- mean(na.omit(as.numeric(liquidity_measures$realizedSpread)))

realized_spread

# Compute the Price Impact.
price_impact <- mean(na.omit(as.numeric(liquidity_measures$priceImpact)))

price_impact
```

## Time Bucket Analysis

Perform analysis on each hour in the dataset.

```{r}
## Split dataset into 24 buckets (24 hours) for analysis.

# amzn_data_list <- list()
# 
# for (i in 0:23)
# {
#   # Construct the time range string.
#   start_hour <- sprintf("%02d:00", i)
#   end_hour <- end_hour <- ifelse(i == 23, "23:59:59", sprintf("%02d:00", i + 1))
#   time_range <- paste("T", start_hour, "/", "T", end_hour, sep = "")
#   
#   # Subset amzn_xts and store in the list.
#   amzn_data_list[[paste("amzn_xts_", i + 1, sep = "")]] <- amzn_data_list[time_range]
# }
# 
# amzn_data_list

amzn_xts_1 <- amzn_xts["T00:00/T01:00"]
amzn_xts_2 <- amzn_xts["T01:00/T02:00"]
amzn_xts_3 <- amzn_xts["T02:00/T03:00"]
amzn_xts_4 <- amzn_xts["T03:00/T04:00"]
amzn_xts_5 <- amzn_xts["T04:00/T05:00"]
amzn_xts_6 <- amzn_xts["T05:00/T06:00"]
amzn_xts_7 <- amzn_xts["T06:00/T07:00"]
amzn_xts_8 <- amzn_xts["T07:00/T08:00"]
amzn_xts_9 <- amzn_xts["T08:00/T09:00"]
amzn_xts_10 <- amzn_xts["T09:00/T10:00"]
amzn_xts_11 <- amzn_xts["T10:00/T11:00"]
amzn_xts_12 <- amzn_xts["T11:00/T12:00"]
amzn_xts_13 <- amzn_xts["T12:00/T13:00"]
amzn_xts_14 <- amzn_xts["T13:00/T14:00"]
amzn_xts_15 <- amzn_xts["T14:00/T15:00"]
amzn_xts_16 <- amzn_xts["T15:00/T16:00"]
amzn_xts_17 <- amzn_xts["T16:00/T17:00"]
amzn_xts_18 <- amzn_xts["T17:00/T18:00"]
amzn_xts_19 <- amzn_xts["T18:00/T19:00"]
amzn_xts_20 <- amzn_xts["T19:00/T20:00"]
amzn_xts_21 <- amzn_xts["T20:00/T21:00"]
amzn_xts_22 <- amzn_xts["T21:00/T22:00"]
amzn_xts_23 <- amzn_xts["T22:00/T23:00"]
amzn_xts_24 <- amzn_xts["T23:00/T23:59:59"]

head(amzn_xts_1)
head(amzn_xts_24)
```

```{r}
## Calculate liquidity measures for each subset.

liquidity_1 <- getLiquidityMeasures(amzn_xts_1)
liquidity_2 <- getLiquidityMeasures(amzn_xts_2)
liquidity_3 <- getLiquidityMeasures(amzn_xts_3)
liquidity_4 <- getLiquidityMeasures(amzn_xts_4)
liquidity_5 <- getLiquidityMeasures(amzn_xts_5)
liquidity_6 <- getLiquidityMeasures(amzn_xts_6)
liquidity_7 <- getLiquidityMeasures(amzn_xts_7)
liquidity_8 <- getLiquidityMeasures(amzn_xts_8)
liquidity_9 <- getLiquidityMeasures(amzn_xts_9)
liquidity_10 <- getLiquidityMeasures(amzn_xts_10)
liquidity_11 <- getLiquidityMeasures(amzn_xts_11)
liquidity_12 <- getLiquidityMeasures(amzn_xts_12)
liquidity_13 <- getLiquidityMeasures(amzn_xts_13)
liquidity_14 <- getLiquidityMeasures(amzn_xts_14)
liquidity_15 <- getLiquidityMeasures(amzn_xts_15)
liquidity_16 <- getLiquidityMeasures(amzn_xts_16)
liquidity_17 <- getLiquidityMeasures(amzn_xts_17)
liquidity_18 <- getLiquidityMeasures(amzn_xts_18)
liquidity_19 <- getLiquidityMeasures(amzn_xts_19)
liquidity_20 <- getLiquidityMeasures(amzn_xts_20)
liquidity_21 <- getLiquidityMeasures(amzn_xts_21)
liquidity_22 <- getLiquidityMeasures(amzn_xts_22)
liquidity_23 <- getLiquidityMeasures(amzn_xts_23)
liquidity_24 <- getLiquidityMeasures(amzn_xts_24)

# # Initialize an empty list to store the liquidity measures.
# liquidity_measures_list <- list()

# # Loop through 24 subsets,
# for (i in 1:24) {
#   # Construct the variable name for each amzn_xts_*
#   xts_var_name <- paste("amzn_xts_", i, sep = "")
# 
#   # Use get() to retrieve the xts object and calculate the liquidity measures.
#   liquidity_measures_list[[paste("liquidity_", i, sep = "")]] <- getLiquidityMeasures(get(xts_var_name))
# }
```

#### Quoted Spread Calculation

```{r}
## Calculate Quoted Spread for each bucket.

# Initialize an empty vector to store the quoted spread means.
quoted_spread_hr <- numeric(24)

# Loop through 24 hours.
for (i in 1:24) {
  # Construct the variable name as a string.
  var_name <- paste("liquidity_", i, sep = "")

  # Use get() to retrieve the dataframe and calculate the mean quoted spread.
  quoted_spread_hr[i] <- mean(as.numeric(get(var_name)$quotedSpread), na.rm = TRUE)
}

quoted_spread_hr
```

#### Quoted Spread Plotting

```{r}
# Plot Quoted Spread for 24 hours.
plot(1:24, quoted_spread_hr, type="p", pch=20, col="blue", main="Quoted Spread", 
     xlab = "Hour", ylab = "Values", ylim = c(0,0.15))
```

#### Effective Spread Calculation

```{r}
## Calculate Effective Spread for each bucket.

# Initialize an empty vector to store the effective spread means.
effective_spread_hr <- numeric(24)

# Loop through 24 hours.
for (i in 1:24) {
  # Construct the variable name as a string.
  var_name <- paste("liquidity_", i, sep = "")

  # Use get() to retrieve the dataframe and calculate the mean quoted spread.
  effective_spread_hr[i] <- mean(as.numeric(get(var_name)$effectiveSpread), na.rm = TRUE)
}

effective_spread_hr
```

#### Effective Spread Plotting

```{r}
# Plot Quoted Spread for 24 hours.
plot(1:24, effective_spread_hr, type="p", pch=20, col="blue", main="Effective Spread", 
     xlab = "Hour", ylab = "Values", ylim = c(0,0.15))
```
